### 2022-04-20 README.Rmd
update readme

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "Robert C. Dunnell’s 'Systematics in Prehistory' at 50"
```

# systematicsinprehistory

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/benmarwick/systematicsinprehistory/master?urlpath=rstudio)  [![.github/workflows/run-on-docker.yaml](https://github.com/benmarwick/systematicsinprehistory/actions/workflows/run-on-docker.yaml/badge.svg)](https://github.com/benmarwick/systematicsinprehistory/actions/workflows/run-on-docker.yaml)

This repository contains the data and code for our paper:

> Authors, (2022). `r Title`. _Evolutionary Human Sciences_  <https://doi.org/xxx/xxx>

Our pre-print is online here:

> Authors, (2022). `r Title`. SocArxiv, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to cite

Please cite this compendium as:

> Authors, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

## Contents

The most important parts of this compendium are:

-   [:dart: \_targets.R](_targets.R): workflow instructions and
    information indicating the order that code needs to be run to
    generate the results. Run `targets::tar_make()` at the R console to
  start the analysis workflow.    
  - [:file\_folder: analysis/paper](/analysis/paper): R Markdown source document
    for manuscript, and R script files. Includes code to reproduce the figures and
    tables generated by the analysis. It also has a rendered version,
    `paper.docx`, suitable for reading (the code is replaced by figures
    and tables in this file)
  - [:file\_folder: analysis/data](/analysis/data): Data used in the analysis.
  - [:file\_folder: analysis/figures](/analysis/figures): Plots and other
    illustrations

## How to run in your broswer or download and run locally

This research compendium has been developed using the statistical programming
language R. To work with the compendium, you will need
installed on your computer the [R software](https://cloud.r-project.org/)
itself and optionally [RStudio Desktop](https://rstudio.com/products/rstudio/download/).

The simplest way to explore the text, code and data is to click on
[binder](https://mybinder.org/v2/gh/benmarwick/systematicsinprehistory/master?urlpath=rstudio)
to open an instance of RStudio in your browser, which will have the
compendium files ready to work with. Binder uses rocker-project.org
Docker images to ensure a consistent and reproducible computational
environment. These Docker images can also be used locally.

You can download the compendium as a zip from from this URL:
[master.zip](/archive/master.zip). After unzipping you can run our code by followinfg these steps:

1. open the `.Rproj` file in RStudio      
2. run `renv::restore()` to download and install the packages required to run our code     
3. run `targets::tar_make()` to run our analysis R code in the `analysis` directory in order, the final step of this will be 
 knitting the R Markdown document to produce our submitted manuscript, `paper.docx`

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.